The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-12 rational.tex
_
explain the semantics

We are able to do this because in {{langRevEE}} every morphism is an
isomorphism and hence has an inverse morphism, whereas in compact
closed categories with no underlying dagger structure we can only
construct an op-functor.

@@

%subcode{opsem} include main
%! columnStyle = rc
%   & b1 * b2
% = & -(-(b1*b2)) 
% = & -((-b1)*b2)
% = & (-b1)*(-b2)

@@

@@

@@

\item To add negative types we add the following rules to the
  reductions above:

  \begin{enumerate}
  \item

%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% <eps+, v, C, s>_{fwd} |--> <eps+, left (-v'), C, s[v <><> right v']>_{bck}
% <eps+, v, C, s>_{fwd} |--> <eps+, right v', C, s[v <><> left (-v')]>_{bck}

    \item
%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% <eta+, v, C, s>_{bck} |--> <eta+, left (-v'), C, s[v <><> right v']>_{fwd}
% <eta+, v, C, s>_{bck} |--> <eta+, right v', C, s[v <><> left (-v')]>_{fwd}

@@

The additions formalize our previous discussions and should not be surprising
at this point.

>>>>

The {{inv}} construction is similar.  We are able to do this because
in {{langRevEE}} every morphism is an isomorphism and hence has an
inverse morphism, whereas in compact closed categories with no
underlying dagger structure we can only construct an op-functor.

@@

%% %subcode{opsem} include main
%% %! columnStyle = rc
%% %   & b1 * b2
%% % = & -(-(b1*b2)) 
%% % = & -((-b1)*b2)
%% % = & (-b1)*(-b2)

{{b1 * b2 = -(-(b1*b2)) = -((-b1)*b2) = (-b1)*(-b2)}}

@@

As explained previously {{eta*}} creates two values {{1/v}} and {{1/v}}
from a single logic variable.

@@

Similarly, {{eps*}} unifies the values on incoming wires. The incoming
value {{v}} represents the values of both wires and hence unifying
them is accomplished in terms of an intermediate logic variable
{{v'}}.

@@

\item To add negative types we add the following rules to the
  reductions above. The additions formalize our previous discussions
  and should not be surprising at this point.

  \begin{enumerate}
  \item The rules for {{eps+}} essentially transfer control from the
    forward evaluator (whose states are tagged by $\triangleright$) to
    the backward evaluator (whose states are tagged by
    $\triangleleft$). In other words, after an {{eps+}} the direction
    of the world is reversed. The pattern matching done by the
    unification ensures that a value on the {{right}} wire is tagged
    to be negative and transfered to the {{left}} wire, and vice versa.
%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% <eps+, v, C, s>_{fwd} |--> <eps+, left (-v'), C, s[v <><> right v']>_{bck}
% <eps+, v, C, s>_{fwd} |--> <eps+, right v', C, s[v <><> left (-v')]>_{bck}
   %
    Note that there is no evaluation rule for {{eta+}} in the forward
    evaluator. This corresponds to the fact that there is no value of
    type {{0}} and hence the forward evaluator can never execute an
    {{eta+}}.

    \item The rules for {{eta+}} are added to backward evaluator. A
      program executing backward starts executing forward after the
      execution of the {{eta+}}. Dual to the previous case, there is
      no rule for {{eps+}} in the backward evaluator since the output
      type of {{eps+}} is {{0}}.
%subcode{opsem} include main
%! columnStyle = rclr
%! fwd = \triangleright
%! bck = \triangleleft
% <eta+, v, C, s>_{bck} |--> <eta+, left (-v'), C, s[v <><> right v']>_{fwd}
% <eta+, v, C, s>_{bck} |--> <eta+, right v', C, s[v <><> left (-v')]>_{fwd}

@@
